---
title: "More Modeling Basics"
author: "Ranae Dietzel and Andee Kaplan"
date: "November 28, 2016"
ratio: 16x10
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    theme: ribbon
---

##More modeling as a tool for exploration  


##Interactions  

```{r setup, message = FALSE, echo=FALSE}
library(tidyverse)
library(nycflights13)
library(lubridate)

library(modelr)
options(na.action = na.warn)
```

```{r, fig.width=6, fig.height=4}
ggplot(sim3, aes(x1, y)) + 
  geom_point(aes(colour = x2))
```

##  
There are two possible models you could fit to this data:

```{r}
mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3)
```

When you add variables with `+`, the model will estimate each effect independent of all the others. It's possible to fit the so-called interaction by using `*`. For example, `y ~ x1 * x2` is translated to `y = a_0 + a_1 * a1 + a_2 * a2 + a_12 * a1 * a2`. Note that whenever you use `*`, both the interaction and the individual components are included in the model.

##  

To visualise these models we need two new tricks:

1.  We have two predictors, so we need to give `data_grid()` both variables. 
    It finds all the unique values of `x1` and `x2` and then generates all
    combinations.  
1.  To generate predictions from both models simultaneously, we can use 
    `gather_predictions()` which adds each prediction as a row. The
    complement of `gather_predictions()` is `spread_predictions()` which adds 
    each prediction to a new column.  
    
## 
Together this gives us:

```{r}
grid <- sim3 %>% 
  data_grid(x1, x2) %>% 
  gather_predictions(mod1, mod2)
grid
```  

##  
We can visualise the results for both models on one plot using facetting:

```{r, fig.width=6, fig.height=4}
ggplot(sim3, aes(x1, y, colour = x2)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred)) + 
  facet_wrap(~ model)
```

##  
Which model is better for this data? We can take look at the residuals. 

```{r, fig.width=6, fig.height=4}
sim3 <- sim3 %>% 
  gather_residuals(mod1, mod2)

ggplot(sim3, aes(x1, resid, colour = x2)) + 
  geom_point() + 
  facet_grid(model ~ x2)
```

##Why are low quality diamonds more expensive?  
```{r}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
```  

##  

```{r}
ggplot(diamonds, aes(color, price)) + geom_boxplot()
```  

##  

```{r}
ggplot(diamonds, aes(clarity, price)) + geom_boxplot()
```

##  
```{r}
ggplot(diamonds, aes(carat, price)) + 
  geom_point()
```  

##  

```{r}
diamonds2 <- diamonds %>% 
  filter(carat <= 2.5) %>% 
  mutate(lprice = log2(price), lcarat = log2(carat))
```

```{r, echo=FALSE, width=6, height=4}
ggplot(diamonds2, aes(lcarat, lprice)) + 
  geom_point()
```
##Identify linear pattern so we can remove it  
```{r}
mod_diamond <- lm(lprice ~ lcarat, data = diamonds2)
```  

##  
```{r}
grid <- diamonds2 %>% 
  data_grid(carat = seq_range(carat, 20)) %>% 
  mutate(lcarat = log2(carat)) %>% 
  add_predictions(mod_diamond, "lprice") %>% 
  mutate(price = 2 ^ lprice)  
```  

```{r, echo=FALSE}
grid  
```  

##  

```{r, fig.width=6, fig.height=4}
ggplot(diamonds2, aes(carat, price)) + 
  geom_point() + 
  geom_line(data = grid, colour = "red", size = 1)
```  

##  

```{r}
diamonds2 <- diamonds2 %>% 
  add_residuals(mod_diamond, "lresid")

ggplot(diamonds2, aes(lcarat, lresid)) + 
  geom_point()
```  

## Cut 

```{r}
ggplot(diamonds2, aes(cut, lresid)) + geom_boxplot()
```   

## Color 
```{r}
ggplot(diamonds2, aes(color, lresid)) + geom_boxplot()  
```  

## Clarity 
```{r}
ggplot(diamonds2, aes(clarity, lresid)) + geom_boxplot()
```  

## A more complicated model

If we wanted to, we could continue to build up our model, moving the effects we've observed into the model to make them explicit. For example, we could include `color`, `cut`, and `clarity` into the model so that we also make explicit the effect of these three categorical variables:

```{r}
mod_diamond2 <- lm(lprice ~ lcarat + color + cut + clarity, data = diamonds2)
```  

##  
```{r}
grid <- diamonds2 %>% 
  data_grid(cut, .model = mod_diamond2) %>% 
  add_predictions(mod_diamond2)
grid
```  

##
```{r}
ggplot(grid, aes(cut, pred)) + 
  geom_point()
```  

## What affects the number of daily flights?  
```{r}
daily <- flights %>% 
  mutate(date = make_date(year, month, day)) %>% 
  group_by(date) %>% 
  summarise(n = n())
daily  
```  

##  
```{r}
ggplot(daily, aes(date, n)) + 
  geom_line()
```  

## Day of week

Understanding the long-term trend is challenging because there's a very strong day-of-week effect that dominates the subtler patterns. Let's start by looking at the distribution of flight numbers by day-of-week:

```{r, echo=FALSE, fig.width=6, fig.height=4}
daily <- daily %>% 
  mutate(wday = wday(date, label = TRUE))
ggplot(daily, aes(wday, n)) + 
  geom_boxplot()
```  

##  

One way to remove this strong pattern is to use a model. First, we fit the model, and display its predictions overlaid on the original data:

```{r, fig.width=4, fig.height=2}
mod <- lm(n ~ wday, data = daily)

grid <- daily %>% 
  data_grid(wday) %>% 
  add_predictions(mod, "n")

ggplot(daily, aes(wday, n)) + 
  geom_boxplot() +
  geom_point(data = grid, colour = "red", size = 4)
```  

##  
Next we compute and visualise the residuals:

```{r, fig.height=4, fig.width=6}
daily <- daily %>% 
  add_residuals(mod)
daily %>% 
  ggplot(aes(date, resid)) + 
  geom_ref_line(h = 0) + 
  geom_line()
```  

##  
Our model seems to fail starting in June: you can still see a strong 
    regular pattern that our model hasn't captured. Drawing a plot with one 
    line for each day of the week makes the cause easier to see:

    ```{r}
    ggplot(daily, aes(date, resid, colour = wday)) + 
      geom_ref_line(h = 0) + 
      geom_line()
    ```  

##  
There are some days with far fewer flights than expected:

    ```{r}
    daily %>% 
      filter(resid < -100)
    ```  

##  
There seems to be some smoother long term trend over the course of a year.
    We can highlight that trend with `geom_smooth()`:

```{r}
    daily %>% 
      ggplot(aes(date, resid)) + 
      geom_ref_line(h = 0) + 
      geom_line(colour = "grey50") + 
      geom_smooth(se = FALSE, span = 0.20)
```  
    

## Seasonal Saturday effect

Let's first tackle our failure to accurately predict the number of flights on Saturday. A good place to start is to go back to the raw numbers, focussing on Saturdays:

```{r, echo=FALSE, fig.width=6, fig.height=4}
daily %>% 
  filter(wday == "Sat") %>% 
  ggplot(aes(date, n)) + 
    geom_point() + 
    geom_line() +
    scale_x_date(NULL, date_breaks = "1 month", date_labels = "%b")
```  

##  

Lets create a "term" variable that roughly captures the three school terms, and check our work with a plot:

```{r}
term <- function(date) {
  cut(date, 
    breaks = ymd(20130101, 20130605, 20130825, 20140101),
    labels = c("spring", "summer", "fall") 
  )
}

daily <- daily %>% 
  mutate(term = term(date)) 
```  

##
```{r, fig.height=4, fig.width=6}
daily %>% 
  filter(wday == "Sat") %>% 
  ggplot(aes(date, n, colour = term)) +
  geom_point(alpha = 1/3) + 
  geom_line() +
  scale_x_date(NULL, date_breaks = "1 month", date_labels = "%b")
```  

##  
It's useful to see how this new variable affects the other days of the week:

```{r, fig.height=4, fig.width=6}
daily %>% 
  ggplot(aes(wday, n, colour = term)) +
    geom_boxplot()
```  

##  

It looks like there is significant variation across the terms, so fitting a separate day of week effect for each term is reasonable. 

```{r}
mod1 <- lm(n ~ wday, data = daily)
mod2 <- lm(n ~ wday * term, data = daily)
```  

##  
This improves our model, but not as much as we might hope:  

```{r}
daily %>% 
  gather_residuals(without_term = mod1, with_term = mod2) %>% 
  ggplot(aes(date, resid, colour = model)) +
    geom_line(alpha = 0.75)
```  

##  

We can see the problem by overlaying the predictions from the model on to the raw data:

```{r, echo=FALSE}
grid <- daily %>% 
  data_grid(wday, term) %>% 
  add_predictions(mod2, "n")

ggplot(daily, aes(wday, n)) +
  geom_boxplot() + 
  geom_point(data = grid, colour = "red") + 
  facet_wrap(~ term)
```  

##
Our model is finding the _mean_ effect, but we have a lot of big outliers, so mean tends to be far away from the typical value. We can alleviate this problem by using a model that is robust to the effect of outliers: `MASS::rlm()`. This greatly reduces the impact of the outliers on our estimates, and gives a model that does a good job of removing the day of week pattern:

```{r, warn = FALSE}
mod3 <- MASS::rlm(n ~ wday * term, data = daily)
```  

##  
```{r}
daily %>% 
  add_residuals(mod3, "resid") %>% 
  ggplot(aes(date, resid)) + 
  geom_hline(yintercept = 0, size = 2, colour = "white") + 
  geom_line()
```   






  



