---
title: "Polishing Figures"
author: "Ranae Dietzel and Andee Kaplan"
ratio: 16x10
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    theme: ribbon
    incremental: true
---
#Polishing Figures  

##Why  
Student: I finished the analysis and look at the great results! What do you think?
```{r, include = FALSE}
#source("script_dataframes.R")
#This script is meant to illustrate subsetting and does not necessarily demostrate the most efficient coding
#practices. Here we import a dataset from National Agricultural Statitistics Service 
#(NASS, https://quickstats.nass.usda.gov/) that focuses on field crops in Iowa.

#Use this to check and set your working directory (or toolbar in RStudio)
getwd()
#setwd()

library(readr) 
crops<-read_csv("NASS-Iowa.csv")  #ctr+enter or command+enter will run the line you are on 

head(crops)  #If you want to glance at the data
tail(crops)  # BTW, hashtag is used to insert commments

#Let's get rid of some columns

  #Choose the columns you want to keep
  fewer_columns<-crops[,c("Year", "State", "Commodity", "Data Item", "Value")]

  #Can also specify by column number  
  notsomany_columns<-crops[,c(2, 6, 16 ,17, 20)]

#Let's get rid of some rows - like everything, there are a lot of ways to do the same thing 
  
  recent<-fewer_columns[fewer_columns$Year != 1866, ]
  
  recent<-fewer_columns[fewer_columns$Year %in% c(1867:2015),]  
  
  recent<-fewer_columns[1:6819,]  
  
#Now let's subset soybeans, just bushels per acre  
  beans<-recent[recent$Commodity == "SOYBEANS" & 
          recent$`Data Item` == "SOYBEANS - YIELD, MEASURED IN BU / ACRE",]  
  
#These columns names are driving me crazy, let's rename  
  colnames(beans)<-c("year", "state", "commodity", "measure", "bu_acre")  
  beans<-beans[,-4] #We don't even need this column, let's remove
  
#Add a unit conversion as a new column
  beans$Mg_ha<-beans$bu_acre*60*0.00045*0.40
 
#Fit a linear model to examine the relationship between time and yield
mod<-lm(beans$bu_acre ~ beans$year)

#Look at the summary
summary(mod)

#Look at names of things the model gives us
names(summary(mod))

#What does it mean by "coefficients"?
summary(mod)$coefficients

#Extract the p-value
summary(mod)$coefficients[,4] 

#Extract the R squared
summary(mod)$r.squared

#Here is something that was extracted (R^2) being useful
library(ggplot2)

#ggplot(beans, aes(x=year, y=bu_acre))+
  #geom_point()+
  #geom_smooth(method='lm')+
  #annotate("text", x=1940, y=50, label = summary(mod)$r.squared, size=5, parse=TRUE)


#######################################################
#Now we want to look at small grains as a group
#Subset most of the small grains listed
smlgrains<-fewer_columns[fewer_columns$Commodity %in% c("OATS", "BARLEY", "WHEAT", "RYE") & 
                       fewer_columns$`Data Item` %in% c("OATS - ACRES HARVESTED","BARLEY - ACRES HARVESTED",
                                                       "WHEAT - ACRES HARVESTED","RYE - ACRES HARVESTED"),]

#These data from NASS have a comma in them and we have to remove it by
 #substituting "nothing" in the place of the comma. You will run into random problems like
 #this a lot and you will always get past it. . .somehow 
smlgrains$Value<-as.numeric(gsub(",", "", smlgrains$Value))

#Here is a glimpse of your future with an awesome package (dplyr)
library(dplyr)
total_smlgrains<-group_by(smlgrains, Year)%>%
  summarise(total=sum(Value))
  
#Check it out 
#plot(total_smlgrains$total~total_smlgrains$Year)

#####################################################
#Don't see many small grains these days, but I do see soybeans. I wonder about the 
 #relationship. Let's compare acre amounts

#Back to beans and this time subset "acres harvested"
bean_acres<-recent[recent$Commodity == "SOYBEANS" & 
                recent$`Data Item` == "SOYBEANS - ACRES HARVESTED",] 

#Now let's work to get the "bean_acres" dataframe and "total_smlgrains" dataframe to 
 #look identical so we tack them together for a tidy result
bean_acres<-bean_acres[,c(1,5)] #this is the information we actually need
colnames(bean_acres)<-c("year", "acres")

#Okay, should have left 'Commodity' in there, but now we don't 
  #have to have SOYBEANS shouted at us
bean_acres$crop<-"soybeans" 

#Make matching column names and "crop" label for total_smlgrains  
colnames(total_smlgrains)<- c("year", "acres")
total_smlgrains$crop<-"smlgrains"

#Now, R, bind these rows together (rbind)
both<-rbind(bean_acres, total_smlgrains)

```
```{r, echo=FALSE}
ggplot(both, aes(x=year, y=acres, group=crop, color=crop))+
  geom_point()+
  geom_smooth()
```

##Why  
Advisor: 

* What is this a graph of?    
* The axes are too small.    
* What does smlgrains mean?  
* People won't be able to read the acre values.  
* Why did Excel make the background grey?   
* That legend will never fit in the publication.  
* I hate these colors.  
* This won't work in black and white  

##Three points on polishing  
* `aes`
* `gg`  
* `theme`
* none of the above  

##
```{r}
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3)+
  geom_smooth()
```

## 
```{r}
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")
```

##
```{r}
theme_set(theme_bw())
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")
```

```{r}
library(scales)
theme_set(theme_bw())
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)
```

##
```{r, eval=FALSE}
theme_set(theme_bw())
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)+
labs(x = "Year",y = "Acres Planted") + 
    theme(#panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          #panel.background = element_blank(),
          axis.line = element_line(),
          legend.position=c(.27,.84), legend.title=element_blank(),
		      legend.text = element_text(size=20),
		      legend.key.size=unit(1, "cm"),
		      axis.title.x = element_text(size=22,vjust=-0.5),
          axis.title.y = element_text(size=22,angle=90),
          axis.text.x = element_text(colour="black", size=20),
          axis.text.y = element_text(colour="black", size=20)) 
```

##
```{r, echo=FALSE}
theme_set(theme_bw())
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)+
labs(x = "Year",y = "Acres Planted") + 
    theme(#panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          #panel.background = element_blank(),
          axis.line = element_line(),
          legend.position=c(.27,.84), legend.title=element_blank(),
		      legend.text = element_text(size=20),legend.key.size=unit(1, "cm"),
		      axis.title.x = element_text(size=22,vjust=-0.5),
          axis.title.y = element_text(size=22,angle=90),
          axis.text.x = element_text(colour="black", size=20),
          axis.text.y = element_text(colour="black", size=20)) 
```


## 
```{r}
my_theme<-theme_bw()+
  theme(#panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          #panel.background = element_blank(),
          axis.line = element_line(),
          legend.position=c(.27,.84), legend.title=element_blank(),
		      legend.text = element_text(size=20),legend.key.size=unit(1, "cm"),
		      plot.title = element_text(size = 22, face = "bold"),
		      axis.title.x = element_text(size=22,vjust=-0.5),
          axis.title.y = element_text(size=22,angle=90),
          axis.text.x = element_text(colour="black", size=20),
          axis.text.y = element_text(colour="black", size=20)) 
```  

##
```{r}
theme_set(my_theme)
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)+
labs(x = "Year",y = "Acres Planted")
```

```{r}
theme_set(my_theme)
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)+
labs(x = "Year",y = "Acres Planted")+
  scale_color_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))+
  scale_shape_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))
```

##
```{r, eval=FALSE}
theme_set(my_theme)
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)+
labs(x = "Year",y = "Acres Planted")+
  scale_color_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))+
  scale_shape_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))
```

##
```{r, echo=FALSE}
theme_set(my_theme)
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)+
labs(x = "Year",y = "Acres Planted")+
  scale_color_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))+
  scale_shape_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))
```

##
```{r, echo=FALSE}
theme_set(my_theme)
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)+
labs(x = "Year",y = "Acres Planted")+
  scale_color_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))+
  scale_shape_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))
```

##Many options to save figures
```{r, eval=FALSE}
ggsave("changes.pdf", plot=cropfig)
```
or
```{r, eval=FALSE}
pdf(file="changes.pdf", height=600, width=600)  
cropfig  
dev.off()
```

Can save in many formats -- .png, .tiff, .jpg, etc. 

#Colors 


## Gradients

Qualitative schemes: no more than 7 colors

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=1.5}
library(RColorBrewer)
library(gridExtra)
new_theme_empty <- theme_bw()
new_theme_empty$line <- element_blank()
new_theme_empty$rect <- element_blank()
new_theme_empty$strip.text <- element_blank()
new_theme_empty$axis.text <- element_blank()
new_theme_empty$plot.title <- element_blank()
new_theme_empty$axis.title <- element_blank()
new_theme_empty$plot.margin <- structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit")
data <- data.frame(x = 1:7, 
                   blues = brewer.pal(7, "Blues"), 
                   set1 = brewer.pal(7, "Set1"), 
                   diverge = brewer.pal(7,"RdBu"))

qplot(xmin = x-.5, xmax = x+.5, ymin = 0, ymax = 1, data = data, geom = "rect", color = I("black"), fill = set1) + 
    scale_fill_identity() + 
    ylab("") + 
    xlab("") + 
    theme(axis.text = element_blank(), 
          axis.ticks = element_blank(), 
          rect = element_blank()) + 
    coord_fixed(ratio = 1) + 
    new_theme_empty
```

Quantitative schemes: use color gradient with only one hue for positive values

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=1.5}
qplot(xmin = x-.5, xmax = x+.5, ymin = 0, ymax = 1, data = data, geom = "rect", color = I("black"), fill = blues) + 
    scale_fill_identity() + 
    ylab("") + 
    xlab("") + 
    theme(axis.text = element_blank(), 
          axis.ticks = element_blank(), 
          rect = element_blank()) + 
    coord_fixed(ratio = 1) + 
    new_theme_empty
```

## More Gradients

Quantitative schemes: use color gradient with two hues for positive and negative values. Gradient should go through a light, neutral color (white)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=1.5}
qplot(xmin = x-.5, xmax = x+.5, ymin = 0, ymax = 1, data = data, geom = "rect", color = I("black"), fill = diverge) + 
    scale_fill_identity() + 
    ylab("") + 
    xlab("") + 
    theme(axis.text = element_blank(), 
          axis.ticks = element_blank(), 
          rect = element_blank()) + 
    coord_fixed(ratio = 1) + 
    new_theme_empty
```

Small objects or thin lines need more contrast than larger areas

## RColorBrewer

R package based on [Cynthia Brewer's color schemes](http://www.colorbrewer2.org)
```{r, echo=FALSE, fig.width=7, fig.height=6}
display.brewer.all()
```

## Color in `ggplot2`

- Factor variable:
    - `scale_colour_discrete`
    - `scale_colour_brewer(palette = ...)`
- Continuous variable:
    - `scale_colour_gradient` (define low, high values)
    - `scale_colour_gradient2` (define low, mid, and high values)
    - Equivalents for fill: `scale_fill_...`

## 

```{r, eval=FALSE}
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)+
labs(x = "Year",y = "Acres Planted")+
  scale_color_brewer(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"), palette="Set1")+
  scale_shape_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))
```

## 
```{r, echo=FALSE}
ggplot(both, aes(x=year, y=acres, group=crop, color=crop, shape=crop))+
  geom_point(size=3, alpha = .6)+
  geom_smooth()+
  ggtitle("Acres of small grains and soybeans in Iowa")+
  scale_y_continuous(labels = comma)+
labs(x = "Year",y = "Acres Planted")+
  scale_color_brewer(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"), palette="Set1")+
  scale_shape_discrete(name="Crop",
                         breaks=c("smlgrains", "soybeans"),
                         labels=c("Small Grains", "Soybeans"))
```

##Great places to look up what you need (all included by Google)  
* [ggplot2 documentation](http://docs.ggplot2.org/current/)
* [Beautiful plotting in R](http://zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/)
* [R Graphics Cookbook](http://www.cookbook-r.com/Graphs/)