---
title: "Your Turn Solution"
author: "Andee Kaplan"
date: "10/7/2016"
output: html_document
---

This is the solution to the "Your Turn" in class from October 7. Below is my function that accomplishes the following:

1. Computes the mean of an input vector.
2. Has error checking to ensure the input vector is numeric or logical, displays an error if not satisfied (see “stop”). 
3. Has a parameter to allow the user to specify if they want to exclude NA values.

```{r setup, include=FALSE}
mymean <- function(x, na.rm = FALSE) {
  # check for numeric or logical input, message a friendly error to user
  if (!(is.numeric(x) | is.logical(x))) {
    stop("Input vector must be numeric or logical!")
  }
  
  # if the user elects to remove NAs, do so.
  if(na.rm) {
    x <- x[!is.na(x)]
  }
  sum(x)/length(x)
}
```

Now we can loop over the columns of the `diamonds` data set and apply my function to all of the numeric columns.

```{r}
library(ggplot2) #diamonds data
for(col in colnames(diamonds)) {
  dat <- diamonds[, col]
  if(all(class(dat) %in% c("numeric", "logical"))) { # sometimes class returns a vector of length > 1
    print(paste(col, ":", mymean(dat)))
  }
}

```

A potential problem with my function is if a user inputs a vector of length 0, the following occurs:

```{r}
 mymean(numeric(0))
```
 This returns a value of "Not a Number", because we have divided by 0. While technically true, it might be nice to add some error handling to give the user a warning when this occurs.